\section{Methods}

We sequenced one wild individual each of Orthochromis sp. ‘red cheek’, Serranochromis sp. ‘checkerboard’, Paralabidochromis paucidens, Harpagochromis vittatus, Paralabidochromis flavus and Harpagochromis sp. ‘checkmate’. We used PCR-free methods (Kozarewa et al. 2009) to generate paired-end reads using Illumina HiSeq. We generated 50–65 million reads per fish – approximately 10-fold coverage – with the exception of Harpagochromis sp. ‘checkmate’, which was sequenced at higher depth for a different project (Neches et al. in preperation). We randomly reduced the number of raw Harpagochromis sp. ‘checkmate’ reads to the level of the other five species. We then aligned all six to the Oreochromis niloticus (oreNil2) genome (Brawand et al. 2014) using bowtie2 (Langmead & Salzberg 2012) and samtools (Li et al. 2009). We then called all confident sites with GATK unifiedgenotyper (McKenna et al. 2010).

Assembling the mitogenome and building a mitogenome tree

In order to establish phylogenetic relationships among the six taxa and to provide an estimate of genetic divergence that is independent of our nuclear genomic data, we extracted and assembled mitogenome data from the reads of each species utilizing the mitobim (Hahn et al. 2013) pipeline and the Oreochromis niloticus mitogenome. We then aligned reads with muscle (Edgar 2004), and manually trimmed each mitogenome to a size of 16 669 bp so that all species had sequence of the same base pair length. We compared several substitution models using jmodeltest 2.1.7 (Darriba et al. 2012), finding that a GTR+G was the best fit to the data. We then used beast 2.2.1 (Bouckaert et al. 2014) with default parameters to infer a phylogenetic tree for our six mitochondrial genomes, using tracer v1.8 (Rambaut & Drummond 2003) to assess performance and treeannotator with 25\% burn-in to produce a consensus mitogenome tree (Fig. 1). Because of the controversy associated with appropriate calibrations of the cichlid molecular clock (Genner et al. 2007; Azuma et al. 2008; Friedman et al. 2013; Schwarzer et al. 2015), we compare only relative ages here.
Genomewide levels of divergence

We then calculated patterns of coverage, called SNPs, df (fixed divergence) and $D_{XY}$ within 10-kb blocks across all the 22 linkage groups of our reference for each lake pair. We used custom r and python scripts, all of which are compiled into an Jupyter notebook (Perez & Granger 2007). 10-kb blocks where only <80\% of all bases were sequenced for at any one of the fish were excluded from further analyses, leaving 39 207 10-kb blocks, over 66\% of the sequence of the entire genome.

df and $D_{XY}$ values were calculated between the ecologically divergent species of each lake, and compared in several ways. A given site was assessed as fixed divergent when both individuals of a given lake are fixed for alternative alleles, but because we only use one individual per species, some sites recorded as divergent would be classified as variable if several individuals within a species were used. Therefore, our estimates of fixed divergence are merely an upper bound on the true number of strongly divergent sites.

We first compared overall patterns of fixed divergence between Victoria, Kivu and Mweru with a Kruskal–Wallis test (Fig. 1, Table S1, Supporting information). We then compared whether genomewide fixed divergence in a block ever exceeds the mean plus two times the standard deviation for each lake pair (Table S1, Supporting information). We also compared fixed divergence in outlier regions from the two younger ecomorph pairs to genomewide divergence in the older ecomorph pair using a Kruskal–Wallis test. We compare the top 1\%, 0.1\% and 0.01\% outliers (Table S1, Supporting information).

Establishing whether outlier regions for relative measures of differentiation, like df, also exhibit overall higher absolute sequence divergence is critically important in establishing whether divergent selection in the face of gene flow was likely to have played a role in speciation and species divergence (Charlesworth 1998; Cruickshank & Hahn 2014). Therefore, we compared absolute divergence $D_{XY}$ between regions with elevated df vs the rest of the genome for multiple outlier levels (Fig. 3, Table S2, Supporting information). In order to test whether parallel phenotypic divergent evolution is associated with a signature of parallel genomic divergent evolution between the three species pairs from different lakes, we also compared whether regions exhibiting high df levels in a given lake pair also exhibited high levels of $D_{XY}$ in the other pairs (Fig. 4, Table S2, Supporting information).
Results

Our time-calibrated mitogenome tree (Fig. 1) reveals that the haplotypes representing our sympatric species pairs are reciprocally monophyletic, with large differences in the age of the Victoria and Kivu species pairs relative to the Mweru species pair, with the 100 best supported trees all indicating far younger divergence in Lake Victoria and Kivu. The inferred split between the two Kivu species is older than that inferred for the two Victoria species, but the 95\% HPD estimates of node height overlap for the two lakes. This indicates that it is difficult to confidently assert an older age for Kivu divergence from mitogenome data alone, although the patterns of fixed divergence across the genome we discuss below also suggest an older age for Kivu. Divergence times in both Kivu and Victoria are <5\% of the inferred divergence time for the two Mweru species, supporting our designation of old and young radiations.

Patterns of fixed divergence in the nuclear genome show a similar pattern, with Victoria exhibiting significantly less fixed divergence (df) than Kivu (P < 0.001, Table S1, Supporting information). Both Victoria and Kivu show much less divergence than Mweru (Fig. 2, Table S1, Supporting information). The distributions of df suggest that what has mainly changed from the young pairs to the old pair is a rise of the genomewide background differentiation, indicating a sustained period of complete reproductive isolation in the Mweru pair. The two young pairs have nearly twice the percentage of genome blocks above the df threshold than the older pair (Table S1, Supporting information). The 1\% and 0.1\% df outliers in Victoria and Kivu are lower than df in Mweru (both comparisons P < 0.01), but the very small number of outliers at the 0.01\% level is not distinguishable (Table S1, Supporting information).

Figure 2. 
Figure 2. Open in figure viewerDownload Powerpoint slide
(A, B) Distribution of df values in 10-kb blocks across the genome. Block heights are scaled to the maximum df value in Lake Mweru. (C). Genomewide df values per 10-kb block, sorted in descending order.
Comparison of outlier regions in the young Victoria and Kivu pairs revealed lower divergence than the average genomewide fixed divergence in Mweru, regardless of the outlier threshold applied (all comparisons P < 0.01, Table S2, Supporting information). For each species pair and for all outlier percentage levels, all df outlier blocks also exhibited increased levels of $D_{XY}$ compared to non-df-outlier blocks (all comparisons P < 0.0001; Fig. 3). df outlier blocks predicted in the Lake Victoria species pair at 1\% and 0.1\% outlier thresholds, also exhibited increased levels of $D_{XY}$ in both the Lake Kivu and the Lake Mweru pair (all comparisons P < 0.001), but this was not true at the 0.01\% outlier level (P > 0.05 for all). A similar pattern was observed for Kivu blocks in Victoria and Mweru, as well as for Mweru blocks in Kivu and Victoria (Fig. 4, Table S2).

Figure 3. 
Figure 3. Open in figure viewerDownload Powerpoint slide
Absolute genomic divergence patterns in three evolutionary replicates of pairs of archetypal ecomorphs of African cichlid fish (a small rock-dwelling insect picker versus a large piscivore) in the three lakes. Beanplots of $D_{XY}$ values between the whole-genome sequences excluding the top 1\% df outlier regions and between the 1\% df outlier regions for each species pair. Horizontal lines indicate means, and curves indicate data density. Note the different y-axis scales.
Figure 4. 
Figure 4. Open in figure viewerDownload Powerpoint slide
Modest but significant extent of genomic parallelism in divergence between three evolutionary replicates of pairs of archetypal ecomorphs of African cichlid fish (a small rock-dwelling insect picker versus a large piscivore.) Beanplots of $D_{XY}$ in Kivu (A, F), Victoria (B, C) and Mweru (D, E) in regions predicted as top 1\% df outlier regions in both other lakes, and regions not predicted as outliers in the other lakes. Horizontal lines indicate means, and curves indicate data density.