\begin{refsection}

\chapter{A workflow for genome-wide mapping of archaeal transcription factors with ChIP-seq }

\chapterauthor{Elizabeth G. Wilbanks, David J. Larsen, Russell Y. Neches, Andrew I. Yao, Chia-Ying Wu, Rachel A. S. Kjolby, Marc T. Facciotti}

\section{Author contributions}

The project was conceived and initiated by Elizabeth G. Wilbanks and
Marc T. Facciotti. Experiments were performed by Elizabeth G.
Wilbanks, David J. Larsen Andrew I. Yao, Chia-Ying Wu and Rachel A. S.
Kjolby. Analysis was performed by Elizabeth G. Wilbanks and Russell Y.
Neches. Analytical tools were written by Russell Y. Neches. The
manuscript was written by Elizabeth G. Wilbanks with contributions
from Russell Y. Neches and Marc T. Facciotti and edited by David J.
Larsen and Andrew I. Yao.

% NOTE : I can't remember if Chia-Ying and Rachel worked on the
% manuscript. I don't remember that they did, but I can't say either
% way.

\clearpage

\section{Publication status}

This chapter appears in press under a Creative Commons Attribution
Non-Commercial License.

\begin{quote}
Wilbanks EG, Larsen DJ, Neches RY, Yao AI, Wu CY, Kjolby RA, Facciotti
MT. {\bf A workflow for genome-wide mapping of archaeal transcription
factors with ChIP-seq.} {\em Nucleic acids research.} 2012 Feb 9.
\end{quote}

\subfile{SaltyBugs/abstract}
\subfile{SaltyBugs/introduction}
\subfile{SaltyBugs/methods}
\subfile{SaltyBugs/results}
\subfile{SaltyBugs/discussion}

\section{Accession Numbers}

All high-throughput sequencing data generated in this work are available via \href{http://www.biotorrents.net/details.php?id=259}{\tt BioTorrents} or from \href{http://www.bme.ucdavis.edu/facciotti/resources_data/data/}{our lab website}. The open access Pique package and source code can be obtained via GitGub at \url{https://github.com/ryneches/pique}. Primer and plasmid sequences used in this study are available in Supplementary Information .

\section{Supplementary Data}

Supplementary Data are available at NAR Online: Supplementary Tables 1-4, Supplementary Figures 1-6, Supplementary Information and Supplementary Reference ( 44 ).

\section{Funding}

UC Davis Startup Funds to M.T.F., an NSF GRFP to E.G.W. and DARPA award (HR0011-05-1-0057) to R.Y.N. Funding for open access charge: UC Davis Startup Funds to M.T.F.

Conflict of interest statement : {\em None declared.}

%\printbibliography[heading=subbibliography,prefixnumbers={\thesection-}]
\printbibliography[heading=subbibliography]

\end{refsection}
