\begin{refsection}

\chapter{The Common Oceanographer: Crowdsourcing the Collection of Oceanographic Data}

\chapterauthor{Federico M. Lauro, Svend Jacob Senstius, Jay Cullen, Russell Y. Neches, Rachelle M. Jensen, Mark V. Brown, Aaron E. Darling, Michael Givskov, Diane McDougald, Ron Hoeke, Martin Ostrowski, Gayle K. Philip, Ian T. Paulsen, Joseph J. Grzymski}

\section{Author contributions}

The project was conceived and initiated by Federico M. Lauro. Basic
water chemistry instrumentation was designed, fabricated by Russell Y.
Neches. Firmware for the instrumentation was written by Russell Y.
Neches. Manuscript was written by Federico M. Lauro with contributions
from Russell Y. Neches and reviewed by the other authors.

% NOTE : I've bugged Fede for contributions of other authors

\subfile{IndigoV/introduction}
\subfile{IndigoV/citizenscience}
\subfile{IndigoV/box1}
\subfile{IndigoV/historical}
\subfile{IndigoV/box2}
\subfile{IndigoV/longterm}
\subfile{IndigoV/lessons}

\section{funding}

The authors acknowledge financial support from the School of BABS (University of New South Wales), the ithree Institute (University of Technology Sydney), the Costerton Biofilm Center (University of Copenhagen), the Singapore Centre on Environmental Life Sciences Engineering (Nanyang Technological University), the Desert Research Institute, and the University of Victoria. JTC was supported by the Natural Sciences and Engineering Research Council of Canada. RYN was supported by an Alfred P. Sloan Foundation grant to Jonathan A. Eisen. FML was supported by a fellowship from the Australian Research Council (DE120102610). The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.

\section{Competing interests}

The authors have declared that no competing interests exist.

\section{Acknowledgements}

We thank the many volunteers who sailed S/Y {\em Indigo V}: Trent Goldsack, Joshua (Jeff Goldblum) Goldstein, Ruth McCance, Raymond Pennotti, Lilia Pitacco, Joseph Podvorec, Alain Talbot, and Jasna Zarkovic.

We also thank Xie Chao, Daniela Drautz, Thomas Jeffries, Helder Marcal, Sophie Mazard, Brett Neilan, Iva Neveux, and Rohan Williams for technical assistance and insightful discussions as well as guidance and encouragement. We thank Mike Harris for generously sharing the YOTREPS data.


\printbibliography[heading=subbibliography]

\end{refsection}
