\begin{refsection}
\chapter{Functional biogeography of ocean microbes revealed through non-negative matrix factorization}

\chapterauthor{Xingpeng Jiang, Morgan G. I. Langille, Russell Y. Neches, Marie Elliot, Simon A. Levin, Jonathan A. Eisen, Joshua S. Weitz, Jonathan Dushoff}

\section{Author contributions}

The project was initiated as part of a collaboration within a DARPA funded project. The idea to re-analyze the Global Ocean Survey data was suggested by Jonathan A. Eisen and Jonathan Dushoff, and the statistical approach we ultimately chose was suggested by Joshua S. Weitz and Xingpeng Jiang. Morgan G. I. Langille searched for the predicted protein domains from the PFAM database and generated the matrixes that were used in the analysis. Environmental factors (satellite data, in particular) was investigated by Russell Y. Neches and incorporated into the analysis by Xingpeng Jiang. The analysis was led by Xingpeng Jiang, and contributions were made by all authors during a weekly conference call hosted by Jonathan Dushoff over the course of about one year. The manuscript was written collaboratively on a wiki by all authors, with the major portion of the writing by Xingpeng Jiang.

\section{Publication status}

This chapter appears in press under a Creative Commons Attribution
license.

\begin{quote}
Jiang X, Langille MG, Neches RY, Elliot M, Levin SA, Eisen JA, Weitz
JS, Dushoff J. {\bf Functional biogeography of ocean microbes revealed
through non-negative matrix factorization.} {\em PLoS One.} 2012 Sep
18.
\end{quote}

\subfile{NMF/abstract}
\subfile{NMF/intro}
\subfile{NMF/factorization}
\subfile{NMF/biogeo}
\subfile{NMF/discussion}
\subfile{NMF/methods}
\subfile{NMF/funding}
\subfile{NMF/ack}

\printbibliography[heading=subbibliography]

\subfile{NMF/supp}

\end{refsection}
