\begin{refsection}

\chapter{Growth of 48 built environment bacterial isolates on board the International Space Station (ISS)}

\chapterauthor{David A. Coil, Russell Y. Neches, Jenna M. Lang, Wendy E. Brown, Mark Severance, Darlene Cavalier, Jonathan A. Eisen}

\section{Author contributions}

The project was conceived and initiated by Russell Y. Neches, Mark
Severance, Darlene Cavalier and Jonathan A. Eisen. Fieldwork was
coordinated by Darline Cavalier and performed by David A. Coil,
Russell Y. Neches, Jenna M. Lang, Wendy E. Brown and many other
contributors. Laboratory work, including culturing, isolations, Sanger 
sequencing and identification was performed by Jennifer C. Flanagan,
Ruth D. Lee, Hannah E. Holland-Moritz, Alex Alexiev and Madison I.
Dunitz. The plates that actually flew were prepaired by Ruth D. Lee
and inoculated by Jennifer C. Flanagan and Madision I. Dunitz. 
Spaceflight logistics were coordinated by Russell Y. Neches and Jenna
M. Lang.

The experimental design was conceived by Russell Y. Neches, and the
protocol was developed by David A. Coil, Russell Y. Neches, Jenna M.
Lang. Russell Y. Neches, David A. Coil, Jenna M. Lang, Wendy E. Brown
and Mark Severance adapted and refined the protocol for spaceflight.

Software was written and analysis was performed by Russell Y. Neches.
The manuscript was written by David A. Coil and reviewed by the
other authors.

This entire project would have failed outright without the
intervention of Carl W. Carruthers, who joined the spaceflight
services contractor about halfway through the project. Carl identified
and solved a number of critical problems in the weeks and hours before
launch, while the experiment was aboard the space station, and upon
return of the samples. Any one of these issues would have resulted in
a total loss of the experiment.

% NOTE : I need a list of undergraduates who worked on SpaceBugs

\subfile{SpaceGrowth/abstract}
\subfile{SpaceGrowth/introduction}
\subfile{SpaceGrowth/methods}
\subfile{SpaceGrowth/results}

\printbibliography[heading=subbibliography]

\end{refsection}
