\begin{refsection}

\chapter{Spatially extensive microbial biogeography of the Indian Ocean provides insights into the unique community structure of a pristine coral atoll}

\chapterauthor{Thomas C. Jeffries, Martin Ostrowski, Rohan B. Williams, Chao Xie, Rachelle M. Jensen, Joseph J. Grzymski, Svend Jacob Senstius, Michael Givskov, Ron Hoeke, Gayle K. Philip, Russell Y. Neches, Daniela I. Drautz-Moses, Caroline Ch{\'e}nard, Ian T. Paulsen, Federico M. Lauro}

\section{Author contributions}

The project was conceived and initiated by Federico M. Lauro. Basic
water chemistry instrumentation was designed, fabricated by Russell Y.
Neches. Firmware for the instrumentation was written by Russell Y.
Neches. Manuscript was written by Federico M. Lauro with contributions
from Russell Y. Neches and reviewed by the other authors.

\section{Publication status}

This chapter appears in press under a Creative Commons Attribution 4.0 International License.

\begin{quote}
Jeffries TC, Ostrowski M, Williams RB, Xie C, Jensen RM, Grzymski JJ,
Senstius SJ, Givskov M, Hoeke R, Philip GK, Neches RY. {\bf Spatially
extensive microbial biogeography of the Indian Ocean provides insights
into the unique community structure of a pristine coral atoll.}
{\em Scientific reports.} 2015 Oct 20.
\end{quote}

% NOTE : I've asked Fede to describe the contributions of other
% authors

\subfile{Chagos/abstract}
\subfile{Chagos/introduction}
\subfile{Chagos/results}
\subfile{Chagos/methods}

\section{funding}

The authors acknowledge financial support from the School of BABS (University of New South Wales), the ithree Institute (University of Technology Sydney), the Costerton Biofilm Center (University of Copenhagen), the Singapore Centre on Environmental Life Sciences Engineering (Nanyang Technological University), the Desert Research Institute, and the University of Victoria. JTC was supported by the Natural Sciences and Engineering Research Council of Canada. RYN was supported by an Alfred P. Sloan Foundation grant to Jonathan A. Eisen. FML was supported by a fellowship from the Australian Research Council (DE120102610). The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.

\section{Competing interests}

The authors declare no competing financial interests.

\section{Acknowledgements}

We thank Dr. Aurore Trottet and Dr. Ole Larsen from DHI for the loan of the YSI sonde model 6600V2. We thank Iva Neveux, Chin Lui Wesley Goi and Krithika Arumugam for technical assistance and Dr. Patrick Martin for insightful discussions. We acknowledge the assistance of the staff at the British Indian Ocean Territory administration for granting us permission to sample the waters in the Chagos Archipelago. We gratefully acknowledge the many volunteers and scientists who sailed S/Y {\em Indigo V} during the Indian Ocean passage from Cape Town to Singapore: Trent Goldsack, Joshua (Jeff Goldblum) Goldstein, Ruth McCance, Raymond Pennotti, Mark V. Brown, Diane McDougald, Jay Cullen, Lilia Pitacco, Joseph Podvorec, Alain Talbot and Jasna Zarkovic. This study was financially supported by the School of BABS (University of New South Wales), the ithree and C3 Institutes (University of Technology Sydney), the Costerton Biofilm Center (University of Copenhagen), the Singapore Centre on Environmental Life Sciences Engineering (SCELSE), the Department of Chemistry and Biomolecular Sciences (Macquarie University) and the Desert Research Institute. The research at the Singapore Centre on Environmental Life Sciences Engineering (SCELSE) is supported by the National Research Foundation Singapore, Ministry of Education to Nanyang Technological University and National University of Singapore, under its Research Centre of Excellence Programme. RYN was supported by an Alfred P. Sloan Foundation grant to Jonathan A. Eisen. FML was supported by a fellowship from the Australian Research Council (DE120102610).

\section{Contributions}

F.M.L. conceived the study. M.O., R.M.J., J.J.G., S.J.S., M.G., R.H., G.K.P. and F.M.L. collected and processed the field samples. J.J.G., D.I.D.-M. and R.Y.N. provided reagents, equipment and technical support. T.C.J., M.O., R.B.W., C.X., S.J.S., C.C., I.T.P. and F.M.L. analysed the data. T.C.J., M.O. and F.M.L. drafted the manuscript and all authors discussed the results and edited the manuscript.

\printbibliography[heading=subbibliography]

\end{refsection}
